---
title: "SEAK Chum `rubias` investigation of long-distance PWS strays"
author: "Kyle Shedd"
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())
library(pacman)
pacman::p_load(tidyverse, lubridate, rubias)
source("~/../R/Functions.GCL.R")
username = readLines("~/../R/usr_pw.txt", n = 1)
.password = readLines("~/../R/usr_pw.txt" , n = 2)[[2]]
```

# Objective

The objective of this notebook is to analyze 2 fish from Prospect Creek, 2019 that had otolith marks from Wally Norenburg Hatchery in PWS.

# Load Genotypes

```{r}
CreateLocusControl.GCL(markersuite = "ChumGolden2011_96SNPs", username = username, password = .password)

LOKI2R.GCL(sillyvec = "CMPROSCR19", username = username, .password)

loci <- LocusControl$locusnames %>% 
  sort() # Initial loci object. This will be used by many fuctions.  
```

# Data QA

Should just be these 2 fish, right?
```{r}
silly_n.GCL("CMPROSCR19")
```

Make sure they aren't missing a ton of genotypes
```{r}
miss_loci <-
  RemoveIndMissLoci.GCL(sillyvec = "CMPROSCR19", proportion = 0.8)
```

Great, and they aren't duplicates of eachother, correct?
```{r}
(
  duplicate_check_95 <-
    CheckDupWithinSilly.GCL(
      sillyvec = "CMPROSCR19",
      loci = loci,
      minnonmissing = 0.6,
      quantile = NULL,
      minproportion = 0.95,
      ncores = 4
    )
)
```

Great, moving on :)

# Baseline

Going to use Andy Barclay's chum baseline that goes from North AK Peninsula to Washington State. For a full description of this baseline, go visit:
"V:\Analysis\5_Coastwide\Chum\NPen2WA_Chum_baseline"

```{r}
(
  chum_baseline <-
    read_csv(
      "V:/Analysis/5_Coastwide/Chum/NPen2WA_Chum_baseline/rubias/baseline/NPen2Wa_Chum_227pops_91loci_base.csv",
      col_types = cols(.default = "c")
    )
)
```

Copy the baseline over to our directory for posterity. First make directory structure.
```{r}
dir.make <- function(path) {
  if(!dir.exists(path)) {dir.create(path)}
}

dir.make(path = "../rubias")
dir.make(path = "../rubias/baseline")
dir.make(path = "../rubias/mixture")
dir.make(path = "../rubias/output")
```

```{r}
file.copy(
  from = "V:/Analysis/5_Coastwide/Chum/NPen2WA_Chum_baseline/rubias/baseline/NPen2Wa_Chum_227pops_91loci_base.csv",
  to = "../rubias/baseline/NPen2Wa_Chum_227pops_91loci_base.csv",
  overwrite = TRUE,
  recursive = FALSE,
  copy.date = TRUE
)
```

## Get Baseline Objects
```{r baseline objects}
attach("V:\\Analysis\\5_Coastwide\\Chum\\NPen2WA_Chum_baseline\\NPen2WA_Chum_baseline.Rdata")

loci91 <- loci91
sillyvec227 <- sillyvec227
groupvec19 <- groupvec19
groups19 <- groups19
baseline <- baseline
grcol <- grcol

detach()

source("~/../R/Functions.GCL.R")
```

# Loci

What loci are in the baseline? I think Andy has some combined markers in there...we should sort that out *prior* to creating our mixture!
```{r}
(
  loci_baseline <-
    str_subset(string = names(chum_baseline), pattern = "\\.1") %>%
    str_remove(string = ., pattern = "\\.1")
)
```

```{r}
all.equal(loci_baseline, loci91)
```


Cool, looks like just one combined marker.
```{r}
CombineLoci.GCL(
  sillyvec = "CMPROSCR19",
  markerset = c("Oke_Cr30", "Oke_Cr386", "Oke_ND3-69"),
  update = TRUE,
  delim = "."
)
```


# Mixture Analysis

## Create Mixture
```{r}
(
  CMPROSCR19_mix <-
    create_rubias_mixture(
      sillyvec = "CMPROSCR19",
      loci = loci_baseline,
      path = "../rubias/mixture/"
    )
)
```

## Analyze

### MCMC

Flat prior by group (not pop)
```{r}
(flat_prior <- tibble(
  collection = sillyvec227,
  pi_param = Prior.GCL(groupvec = groupvec19, groupweights = rep(1 /
                                                                   max(groupvec19), max(groupvec19)))
))
```

Run!!!
```{r}
(
  mix.out <-
    run_rubias_mixture(
      reference = chum_baseline,
      mixture = CMPROSCR19_mix,
      group_names = groups19,
      gen_start_col = 5,
      method = "MCMC",
      pi_prior = flat_prior, 
      path = "../rubias/output/"
    )
)
```

Summarize stock composition to RG
```{r}
(prop_est  <-
   custom_combine_rubias_output(rubias_output = mix.out, bias_corr = FALSE))
```

Individual posteriors
```{r}
mix.out$indiv_posteriors %>% 
  group_by(indiv, repunit) %>% 
  summarize(prob = sum(PofZ)) %>% 
  slice_max(prob, n = 5)
```

Hrmm, well it definitely looks like `CMPROSCR19_1090` is from PWS (> 0.95), however, there is a bit more uncertainty around `CMPROSCR19_759` (PWS p ~ 0.80).

### PB

Run!!!
```{r}
dir.make("../rubias/output_PB/")

(
  mix.out_PB <-
    run_rubias_mixture(
      reference = chum_baseline,
      mixture = CMPROSCR19_mix,
      group_names = groups19,
      gen_start_col = 5,
      method = "PB",
      pi_prior = flat_prior, 
      path = "../rubias/output_PB/"
    )
)
```

Summarize stock composition to RG
```{r}
(prop_est_PB  <-
   custom_combine_rubias_output(rubias_output = mix.out_PB, bias_corr = TRUE))
```

Individual posteriors
```{r}
mix.out_PB$indiv_posteriors %>% 
  group_by(indiv, repunit) %>% 
  summarize(prob = sum(PofZ)) %>% 
  slice_max(prob, n = 5)
```

Right, the parametric bootstrap option (bias correction) *only* affects the stock composition estimates, not the individual genotype scaled likelihoods!!!

Hrmm, well it definitely looks like `CMPROSCR19_1090` is from PWS (> 0.95), however, there is a bit more uncertainty around `CMPROSCR19_759` (PWS p ~ 0.80).

# Conclusions

It certainly seems very likely that at least one of these fish is a long-distance stray from PWS (p > 0.95), and the other is most likely from PWS as well (p ~ 0.80.