---
title: "SEAK Chum Spatial Patterns of Hatchery vs. Natural fish"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
while(!require(ggnewscale)){devtools::install_github("eliocamp/ggnewscale")}

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

# Read Data

Read in StreamSpecimens data from *Finsight* to get lat/long data
```{r}
(finsight <- read_csv("../Stream Specimens/StreamSpecimens_SEAK_2013-2019_CSVReport.csv") %>% 
   distinct() %>% 
   filter(!is.na(SampleCell)) %>% 
   unite(key, c("SampleTrayId", "SampleCell"), sep = "_", remove = FALSE)
 )
```

Read in *OceanAK* data to get latest and greatest otolith reads
```{r}
(oceanak <- read_csv("../OceanAK/Salmon Biological Data 2 - Kyle's Filter_20191220.csv") %>% 
   unite(key, c("DNA Tray Code", "DNA Tray Well Code"), sep = "_", remove = FALSE))
```

Crap, *OceanAK* doesn't have the 2019 data, presumably because we don't have the 2019 samples in-house yet, thus they are not in *LOKI*, so maybe the view in *OceanAK* doesn't work?
```{r}
oceanak %>% 
  count(`Silly Code`)
```

Just keep key and Otolith Mark Present
```{r}
(oceanak_mod <- oceanak %>% 
   select(key, `Otolith Mark Present`)
)
```

Ugh, it is missing the 2019 data, WTF! Read in 2019 data from Bev
```{r}
(mta <- read_csv("../OceanAK/data for Kyle 12-20-19.csv"))
```

Make key and modify otolith mark to boolean
```{r}
(mta_mod <- mta %>% 
   mutate(dna_tray_cell_code = as.character(as.numeric(CELL_NUMBER)),
          dna_tray_code = as.numeric(DNA_TRAY_CODE)) %>% 
   unite(key, c("dna_tray_code", "dna_tray_cell_code"), sep = "_", remove = FALSE) %>% 
   mutate("Otolith Mark Present" = case_when(MARK_PRESENT == 0 ~ "NO",
                                             MARK_PRESENT == -1 ~ "YES",
                                             TRUE ~ NA_character_)) %>% 
   select(key, "Otolith Mark Present")
)
```

Join OceanAK (2013-2018) and MTA (2019) otolith data
```{r}
(oto <- bind_rows(oceanak_mod, mta_mod))
```


# Join and Modify Data

Prepare to join *Finsight* and *OceanAK* based on key
```{r}
table(oto$key %in% finsight$key)
table(finsight$key %in% oto$key)
```

Join otolith data with *Finsight*, verify conflicts first
```{r}
finsight %>% 
  left_join(oto, by = "key") %>% 
  count(MarkPresent, `Otolith Mark Present`)
```

Join and create a single variable for origin
```{r}
finsight_mod <- finsight %>% 
  left_join(oto, by = "key") %>% 
  mutate(otolith_mark_present = case_when(MarkPresent == "N" ~ FALSE,
                                          MarkPresent == "Y" ~ TRUE,
                                          MarkPresent == "Null" & `Otolith Mark Present` == "NO" ~ FALSE,
                                          MarkPresent == "Null" & `Otolith Mark Present` == "YES" ~ TRUE,
                                          MarkPresent == "Null" & `Otolith Mark Present` == NA ~ NA,
                                          MarkPresent == NA ~ NA)) %>% 
  mutate(date = mdy(SurveyDate),
         year = year(date),
         DOY = yday(date),
         Longitude = -abs(Longitude)) %>% 
  rename(stream = StreamName) %>% 
  mutate(otolith_read = case_when(otolith_mark_present == NA ~ FALSE,
                                  TRUE ~ TRUE),
         origin = case_when(otolith_mark_present == TRUE ~ "Hatchery",
                            otolith_mark_present == FALSE ~ "Natural",
                            TRUE ~ NA_character_),
         origin = factor(x = origin, levels = c("Natural", "Hatchery")))
```

# Otolith Status
All otoliths read?
```{r}
finsight_mod %>%
  count(year, otolith_read) %>% 
  spread(otolith_read, n, fill = 0)
```

What are otolith reads by year?
```{r}
finsight_mod %>% 
  count(origin, year) %>% 
  spread(origin, n, fill = 0)
```

Calculate raw pHOS
```{r}
finsight_mod %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  count(stream, year, origin) %>% 
  spread(origin, n) %>% 
  group_by(stream, year) %>% 
  summarize(pHOS = round(Hatchery / (Hatchery + Natural), 3) * 100) %>% 
  spread(stream, pHOS)
```

# Sample Sizes
Sample size by year?
```{r}
finsight_mod %>% 
  count(stream, year) %>% 
  spread(stream, n, fill = 0)
```

## Plot all streams and years
```{r fig.width=12}
finsight_mod %>% 
  # filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, fill = origin)) +
  geom_histogram(binwidth = 1) +
  facet_grid(year ~ stream) +
  labs(fill = "Origin") +
  ylab("Number of Samples") +
  xlab("Day of Year") +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

## Plot daily pHOS
```{r warning=FALSE, fig.width = 12}
finsight_mod %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  count(stream, year, DOY, origin) %>% 
  spread(origin, n, fill = 0) %>% 
  group_by(stream, year, DOY) %>% 
  summarize(pHOS = Hatchery / (Hatchery + Natural),
            n = Hatchery + Natural) %>% 
  # filter(stream == "Fish Creek" & year == 2013) %>% 
  ggplot(aes(x = DOY, y = pHOS * 100)) +
  geom_point(aes(size = n), alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, aes(weight = n)) +
  ylim(c(0, 100)) +
  facet_grid(year ~ stream) +
  labs(fill = "Origin") +
  ylab("pHOS") +
  xlab("Day of Year") +
  theme_bw() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

# Location

## Fish Creek
```{r fig.height=10}
finsight_mod %>% 
  filter(stream != "Admiralty Creek" | year != 2019) %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  filter(stream == "Fish Creek") %>% 
  ggplot(aes(x = Longitude, y = Latitude, colour = origin)) +
  geom_jitter(alpha = 0.5, size = 3) +
  coord_quickmap() +
  facet_grid(year ~ stream) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Prospect Creek
```{r fig.width=12}
finsight_mod %>% 
  filter(stream != "Admiralty Creek" | year != 2019) %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  filter(stream == "Prospect Creek") %>% 
  ggplot(aes(x = Longitude, y = Latitude, colour = origin)) +
  geom_jitter(alpha = 0.5, size = 3) +
  coord_quickmap() +
  facet_grid(stream ~ year) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Sawmill Creek
```{r fig.width=12}
finsight_mod %>% 
  filter(stream != "Admiralty Creek" | year != 2019) %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  filter(stream == "Sawmill Creek") %>% 
  ggplot(aes(x = Longitude, y = Latitude, colour = origin)) +
  geom_jitter(alpha = 0.5, size = 3) +
  coord_quickmap() +
  facet_grid(stream ~ year) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Admiralty Creek
```{r fig.height=10}
finsight_mod %>% 
  filter(stream != "Admiralty Creek" | year != 2019) %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  filter(stream == "Admiralty Creek") %>% 
  ggplot(aes(x = Longitude, y = Latitude, colour = origin)) +
  geom_jitter(alpha = 0.5, size = 3) +
  coord_quickmap() +
  facet_grid(year ~ stream) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

# Very Rough Distance

Use some basic trigonometry to calculate distance from mouth on Fish Creek and Prospect Creek, because they are more or less straight.

```{r}
mouth <- finsight_mod %>% 
  filter(stream == "Fish Creek") %>% 
  summarise(Latitude = max(Latitude, na.rm = TRUE),
            Longitude = min(Longitude, na.rm = TRUE))
```

```{r}
finsight_mod %>% 
  filter(stream == "Fish Creek") %>% 
  mutate(distance = sqrt((mouth$Latitude - Latitude)^2 + abs(mouth$Longitude - Longitude)^2)) %>% 
  ggplot(aes(x = distance, fill = origin)) +
  geom_histogram() +
  facet_grid(year ~ stream) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

```{r}
finsight_mod %>% 
  filter(stream == "Fish Creek" & origin %in% c("Natural", "Hatchery")) %>% 
  mutate(distance = sqrt((mouth$Latitude - Latitude)^2 + abs(mouth$Longitude - Longitude)^2)) %>% 
  ggplot(aes(x = distance, fill = origin)) +
  geom_density(alpha = 0.5) +
  facet_grid(year ~ stream) +
  theme_bw() +
  theme(text = element_text(size = 20))
```


## Contour density plots
```{r fig.height=15}
finsight_mod %>% 
  filter(stream == "Fish Creek" & origin %in% c("Natural", "Hatchery")) %>% 
  mutate(distance = sqrt((mouth$Latitude - Latitude)^2 + abs(mouth$Longitude - Longitude)^2)) %>% 
  ggplot(aes(x = DOY, y = distance, colour = origin)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_gradient(low = "grey50", high = "black") +
  facet_grid(year ~ origin) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

# Riverdist

## Fish Creek
```{r}
(riverdist_fish <- read_csv("../GIS/R/fish/fish_distances.csv") %>% 
   rename(Distance = mouthdist) %>% 
   distinct()
)
```

```{r}
(finsight_mod_fish <- finsight_mod %>% 
   unite(Sample, c(SampleTrayId, SampleCell), sep = "_", remove = FALSE) %>% 
   filter(stream == "Fish Creek") %>% 
   select(stream, year, date, DOY, Latitude, Longitude, Sample, SampleTrayId, SampleCell, Sex, MEHLength, SpawningState, Hatchery, Age, origin, otolith_read)
)
```


```{r}
join_fish <- riverdist_fish %>% 
  dplyr::left_join(finsight_mod_fish, by = "Sample")


join_fish  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_gradient(low = "grey50", high = "black") +
  facet_grid(year ~ origin) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Sawmill Creek
```{r}
(riverdist_sawmill <- read_csv("../GIS/R/sawmill/sawmill_distances.csv") %>% 
   rename(Distance = mouthdist) %>% 
   distinct()
)
```

```{r}
(finsight_mod_sawmill <- finsight_mod %>% 
   unite(Sample, c(SampleTrayId, SampleCell), sep = "_", remove = FALSE) %>% 
   filter(stream == "Sawmill Creek") %>% 
   select(stream, year, date, DOY, Latitude, Longitude, Sample, SampleTrayId, SampleCell, Sex, MEHLength, SpawningState, Hatchery, Age, origin, otolith_read)
)
```


```{r, fig.height=10}
join_sawmill <- riverdist_sawmill %>% 
  dplyr::left_join(finsight_mod_sawmill, by = "Sample")


join_sawmill  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_gradient(low = "grey50", high = "black") +
  facet_grid(year ~ origin) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Prospect Creek
```{r}
(riverdist_prospect <- read_csv("../GIS/R/prospect/prospect_distances.csv") %>% 
   rename(Distance = mouthdist) %>% 
   distinct()
)
```

```{r}
(finsight_mod_prospect <- finsight_mod %>% 
   unite(Sample, c(SampleTrayId, SampleCell), sep = "_", remove = FALSE) %>% 
   filter(stream == "Prospect Creek") %>% 
   select(stream, year, date, DOY, Latitude, Longitude, Sample, SampleTrayId, SampleCell, Sex, MEHLength, SpawningState, Hatchery, Age, origin, otolith_read)
)
```


```{r fig.height=10}
join_prospect <- riverdist_prospect %>% 
  dplyr::left_join(finsight_mod_prospect, by = "Sample")


join_prospect  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_gradient(low = "grey50", high = "black") +
  facet_grid(year ~ origin) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Admiralty Creek
```{r}
(riverdist_admiralty <- read_csv("../GIS/R/admiralty/admiralty_distances.csv") %>% 
   rename(Distance = mouthdist) %>% 
   distinct()
)
```

```{r}
(finsight_mod_admiralty <- finsight_mod %>% 
   unite(Sample, c(SampleTrayId, SampleCell), sep = "_", remove = FALSE) %>% 
   filter(stream == "Admiralty Creek") %>% 
   select(stream, year, date, DOY, Latitude, Longitude, Sample, SampleTrayId, SampleCell, Sex, MEHLength, SpawningState, Hatchery, Age, origin, otolith_read)
)
```


```{r fig.height=10}
join_admiralty <- riverdist_admiralty %>% 
  dplyr::left_join(finsight_mod_admiralty, by = "Sample")


join_admiralty  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_gradient(low = "grey50", high = "black") +
  facet_grid(year ~ origin) +
  theme_bw() +
  theme(text = element_text(size = 20))
```

# One plot to rule them all

```{r}
join_fish  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin, group = origin)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.5) +
  scale_fill_gradient(low = "grey50", high = "black") +
  facet_grid(year ~ .) +
  theme_bw() +
  theme(text = element_text(size = 20))


  # new_scale_color() +
  # scale_colour_manual(name = "Origin", values = c("Natural" = origin_colors[1], "Hatchery" = origin_colors[2])) +


gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

origin_colors <- gg_color_hue(2)

join_fish  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin)) + 
  stat_density_2d(geom = "density_2d", show.legend = TRUE) +  # contour lines
  labs(colour = "Origin") +  # change legend title
  guides(colour = guide_legend(override.aes = list(size = 4))) +  # make legend line width larger ~ pseudo fill
  stat_density_2d(data = filter(.data = join_fish, origin == "Natural"), aes(fill = ..level..), geom = "polygon", alpha = 0.5, colour = origin_colors[1], show.legend = FALSE) +  # plot natural
  scale_fill_gradient(low = "white", high = origin_colors[1]) +  # add custom fill colour
  new_scale_fill() +  # switch fill
  stat_density_2d(data = filter(.data = join_fish, origin == "Hatchery"), aes(fill = ..level..), geom = "polygon", alpha = 0.5, colour = origin_colors[2], show.legend = FALSE) +  # add hatchery
  scale_fill_gradient(low = "white", high = origin_colors[2]) +  # add custom fill colour
  facet_grid(year ~ .) +  # facet by year
  theme_bw() +
  theme(text = element_text(size = 20))
```

```{r, fig.width=12}
join_all <- bind_rows(join_fish, join_prospect, join_sawmill)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

origin_colors <- gg_color_hue(2)

join_all  %>% 
  filter(origin %in% c("Natural", "Hatchery")) %>% 
  ggplot(aes(x = DOY, y = Distance, colour = origin)) + 
  stat_density_2d(geom = "density_2d", show.legend = TRUE) +  # contour lines
  labs(colour = "Origin") +  # change legend title
  guides(colour = guide_legend(override.aes = list(size = 4))) +  # make legend line width larger ~ pseudo fill
  stat_density_2d(data = filter(.data = join_all, origin == "Natural"), aes(fill = ..level..), geom = "polygon", alpha = 0.1, colour = origin_colors[1], show.legend = FALSE) +  # plot natural
  scale_fill_gradient2(low = "white", mid = origin_colors[1], high = "darkred") +  # add custom fill colour
  new_scale_fill() +  # switch fill
  stat_density_2d(data = filter(.data = join_all, origin == "Hatchery"), aes(fill = ..level..), geom = "polygon", alpha = 0.1, colour = origin_colors[2], show.legend = FALSE) +  # add hatchery
  scale_fill_gradient2(low = "white", mid = origin_colors[2], high = "darkblue") +  # add custom fill colour
  scale_x_continuous(limits = c(195, 245), breaks = seq(195, 245, by = 20), labels = format(x = (as.Date("2012-12-31") + seq(195, 245, by = 20)), "%b %d")) +  # converty day of year to sample date
  xlab("Sample Date") +
  ylab("Distance from Stream Mouth (m)") +
  facet_grid(stream ~ year, scales = "free_y") +  # facet by year
  theme_bw() +
  theme(text = element_text(size = 20), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

